---
title: "Lab12-Pregnant Smoking"
output: html_document
date: 'April 24, 2022'
---

Note: I need this for Mac: 

```{r Stop the errors!}
r = getOption("repos")
r["CRAN"] = "http://cran.us.r-project.org"
options(repos = r)
```

```{r Packages1}
install.packages("openintro")
```

```{r Packages2}
library(tidyverse) 
library(infer)
library(openintro)
```

```{r Library maybes}
install.packages('tidyr')
install.packages('readr')
```

```{r Library maybes2} 
library(ggplot2)
library(tidyr)
library(readr)
```

Set a seed.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

``{r load NC births data}
data(ncbirths)
help(ncbirths)
``

```{r load NC births data Alternative}
download.file("http://www.openintro.org/stat/data/nc.RData", destfile = "ncbirths")
load("ncbirths")
ncbirths<-nc
```

1. What are the cases in this data set? How many cases are there in our sample?
The cases in this data set are data on individual births recorded in NC. There are 1000 cases.  

1.1. Make summaries of the variables in your dataset, determine which variables are categorical and which are numerical.
fage  <int> numerical
mage <int> numerical
mature <fct>  categorical
weeks <int> numerical
premie <fct>  categorical
visits <int> numerical
marital <fct>  categorical
gained  <int> numerical
weight <dbl> numerical
lowbirthweight <fct> categorical
gender <fct> categorical
habit <fct> categorical
whitemom <fct> categorical
  
-The actual data (integer, factor, numeric) types of your data:
```{r List data type for each variable1}   
str(ncbirths)
```
  

  
1.2. For numerical variables, are there outliers? 
There are incredible outliers for each of these variables. For intance, in this data set, two babies were born weighing one pound.

-Amount (actual data value) and frequency (number of times that value occurs) of each variable:
```{r List data type for each variable2}   
table(ncbirths$fage)
table(ncbirths$mage)
table(ncbirths$weeks)
table(ncbirths$visits)
table(ncbirths$gained)
table(ncbirths$weight)
```

3. Create a filtered data frame called ncbirths_white that contain data only from white mothers. 

```{r Df} 
ncbirths_white<-filter(ncbirths, whitemom == "white")
```

3.1. Calculate the mean of the weights of their babies.
7.25 pounds is the mean weight

```{r Calculate mean} 
summary(ncbirths_white)
```

4. Are the conditions necessary for conducting simulation based inference satisfied? Explain your reasoning.
Yes, because we have a random sample from which we want to estimate a population parameter. We can create a measurement of how much we can expect the sample population to vary from sample to sample. Bootstrapping should get us there. 

5. Run the appropriate hypothesis test. 
The null hypothesis here is that H0:μ=7.43 pounds.

```{r 5 null hypo} 
inference(y = ncbirths$weight, x = ncbirths$habit, est = "mean", type = "ht", null = 0, 
          alternative = "twosided", method = "theoretical")
```


5.1. Visualize the null distribution. 
```{r 5.1 null distro} 
inference(y = ncbirths$weight, x = ncbirths$habit, est = "mean", type = "ci", null = 0, 
          alternative = "twosided", method = "theoretical", 
          order = c("smoker","nonsmoker"))
```

5.2. Calculate the p-value. 
p-value =  0.0184 

5.3. Interpret the results in context of the data and the hypothesis test.
There is a mean difference in average birth weight when comparing smokers versus non-smokers.

Baby weight vs. mother’s age
6. Make side-by-side boxplots displaying the relationship between habit and weight. 

```{r 6}
boxplot(weight~habit,data=ncbirths, main="Relation Between Mother's Habit and Baby's Weight", 
    ylab="Baby's Weight", xlab="Mother Smoker/Non-Smoker")
```

6.1. What does the plot highlight about the relationship between these two variables?
Mothers who smoke have babies of lower weight than mothers who do not smoke. 

7. Before moving forward, save a version of the dataset omitting observations where there are NAs for habit. You can call this version ncbirths_habitgiven.

```{r omit NAs}
ncbirths_habitgiven<- na.omit(ncbirths) 
```


```{r 7.1}
ncbirths_habitgiven %>%
  group_by(habit) %>%
  summarise(mean_weight = mean(weight))
```
  
8. Write the hypotheses for testing if the average weights of babies born to smoking and non-smoking mothers are different.
Null hypothesis: There is no difference in average pounds weight of babies born to smoking mothers versus babies borth to non-smoking mothers.

9. Are the conditions necessary for conducting simulation based inference satisfied? Explain your reasoning.
Yes, the conditions necessary are met. After doing a bit of research, I discovered this answer at: https://rpubs.com/millerfe17/MillerHW5.

```{r 9}
by(ncbirths$weight, ncbirths$habit, length)
```

10. Run the appropriate hypothesis test, calculate the p-value, and interpret the results in context of the data and the hypothesis test.

```{r 10}
inference(y = ncbirths$weight, x = ncbirths$habit, est = "mean", type = "ht", null = 0, alternative = "twosided", method = "theoretical")
```


11. Construct a 95% confidence interval for the difference between the average weights of babies born to smoking and non-smoking mothers.

```{r 11}  
inference(y = nc$weight, x = nc$habit, est = "mean", type = "ci", null = 0, 
          alternative = "twosided", method = "theoretical", 
          order = c("smoker","nonsmoker"))
```
  
Baby weight vs. mother’s age

Focus on two variables: maturemom, lowbirthweight.

12. First, a non-inference task: Determine the age cutoff for younger and mature mothers. Use a method of your choice, and explain how your method works.
I determined that the minimum age of a mature mom is 35. Thus, younger mothers are those aged from the youngest in our sample up to 34. 

```{r showing age cutoff}
library(dplyr)
ncbirths %>% group_by(mature) %>% summarize(min(mage))
```

The other variable of interest is lowbirthweight.

13. Conduct a hypothesis test evaluating whether the proportion of low birth weight babies is higher for mature mothers. State the hypotheses, verify the conditions, run the test and calculate the p-value, and state your conclusion in context of the research question. Use α=0.05

Null hypothesis: There is no difference in average weight of babies between mature mothers and non-mature mothers.
p-value =  0.8526 

```{r 13}
inference(y = ncbirths$weight, x = ncbirths$mature, est = "mean", type = "ht", null = 0, 
          alternative = "twosided", method = "theoretical") 
```

13.1. If you find a significant difference, construct a confidence interval, at the equivalent level to the hypothesis test, for the difference between the proportions of low birth weight babies between mature and younger moms, and interpret this interval in context of the data.
  
```{r 13.1}  
inference(y = ncbirths$weight, x = ncbirths$mature, est = "mean", type = "ci", null = 0, 
          alternative = "twosided", method = "theoretical")   
```  


